---
title: "Text Mining in R --  Ngrams and Log Odds"
author: "Gosse Bouma, Defne Abur"
date: "Feb, 2025"
output:
  pdf_document: default
  html_document: default
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, dpi=150)
library(tidyverse)
library(readr)
library(tidytext)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(tidylo)
library(here)
library(tidytext)
library(gutenbergr)
```

## Goal

The goal of this assignment is:

-   to learn how to collect statistics for combinations of 2 or more
    words (n-grams)
-   to inspect some of the results for negation
-   to experiment with the log-odds method for detecting keywords

## Book Review Dataset Again

We will use the same type of data as for the previous assignment; i.e.,
book reviews from the the [Amazon Review
dataset](https://jmcauley.ucsd.edu/data/amazon/)

We will use the same type of data as for the previous assignment. For
this assignment, however, the samples are 10 times larger (each file
contains 5000 reviews). This time, please pick the sample that
corresponds to the first number of your student number (i.e. student
s123456 will work with pos1.txt and neg1.txt).

All data can be found on [Google
drive](https://drive.google.com/drive/folders/1IEZHHWebxlLLJLYeRxnplFSWFhdwU1iG?usp=sharing)

```{r}
# Load data 
# My snumber is s5788668, so I will use pos5.txt and neg5.txt
pos <- read_lines(here("data", "raw", "corpusComparision", "pos5.txt"))
neg <- read_lines(here("data", "raw", "corpusComparision", "neg5.txt"))
head(pos, 2)
head(neg, 2)
```

This goal of this assignment is to investigate:

-   differences between negative vs positive reviews
-   bigrams
-   negation
-   aspects
-   collocations

In sentiment analysis, aspects refer to a particular aspect of a product
on which a sentiment is expressed. Concretely, a book review may say
"the plot was great, but the writing style was terrible"; "plot" and
"writing style" are aspects.

## A: Tokenizing N-grams

(2 points)

The `readr` package provides a function `read_lines()` that can be used
to read arbitrary text and store the result as a vector of strings. 1.
Use this to read the two files. 2. Next, create a `tibble` dataframe for
the data of the positive review. 3. Create a `tibble` dataframe for the
data of the negative review. 4. Tokenize the review text into bigrams,
as described in [chapter
4.1](http://tidytextmining.com/ngrams.html#tokenizing-by-n-gram) of Text
mining in R.

```{r}
# Create tibble dataframes
pos_df <- tibble(text = pos)
neg_df <- tibble(text = neg)

# Tokenize the review text into bigrams
pos_bigram <- pos_df |> 
  unnest_tokens(bigram, text, token = "ngrams", n = 2) |> 
  filter(!is.na(bigram)) # I understand but disapprove it 
# Adding <s></s> can solve the problem, no need to sacrifice semantic info

neg_bigram <- neg_df |>
  unnest_tokens(bigram, text, token = "ngrams", n = 2) |> 
  filter(!is.na(bigram)) 
```

## B: Splitting N-grams

(3 points)

1.  Split the bigram column into two columns word1 and word2, as
    described in chapter 4.1 (e.g., using the `separate()` function).Do
    not apply stopword filtering.
2.  Determine the 5 most frequent words preceding ‘style’ in positive
    reviews (if there are ties, just report the first 5!)
3.  Determine the 5 most frequent words preceding ‘style’ in negative
    reviews
4.  Repeat the above two steps for the word ‘characters’ (for both
    positive and negative reviews)
5.  Do you see interesting differences between positive and negative
    reviews for these two words? Comment on your findings.
6.  How often do positive and negative reviews mention each of these
    words (i.e., what is the relative frequency of these words)? Comment
    on your findings.

```{r}
# To improve the readability of the code, I will keep it in one chunk, with 
# comments for each step

# B1: Split the bigram column into two columns word1 and word2
pos_bigrams_separated <- pos_bigram %>%
  separate(bigram, c("word1", "word2"), sep = " ")

neg_bigrams_separated <- neg_bigram %>%
  separate(bigram, c("word1", "word2"), sep = " ")

# B2: 5 most frequent words preceding 'style' in positive reviews
pos_style <- pos_bigrams_separated %>%
  filter(word2 == "style") %>%
  count(word1, sort = TRUE) %>%
  head(5)

# B3: 5 most frequent words preceding 'style' in negative reviews
neg_style <- neg_bigrams_separated %>%
  filter(word2 == "style") %>%
  count(word1, sort = TRUE) %>%
  head(5)

# B4: 5 most frequent words preceding 'characters' in pos & neg reviews
pos_characters <- pos_bigrams_separated %>%
  filter(word2 == "characters") %>%
  count(word1, sort = TRUE) %>%
  head(5)

neg_characters <- neg_bigrams_separated %>%
  filter(word2 == "characters") %>%
  count(word1, sort = TRUE) %>%
  head(5)

# Print results
print("Words preceding 'style' in positive reviews:")
print(pos_style)

print("Words preceding 'style' in negative reviews:")
print(neg_style)

print("Words preceding 'characters' in positive reviews:")
print(pos_characters)

print("Words preceding 'characters' in negative reviews:")
print(neg_characters)

# B6: Relative frequency of 'style' and 'characters' in both pos & neg reviews
style_freq <- bind_rows(
  pos_bigrams_separated %>% 
    filter(word2 == "style") %>%
    summarize(type = "positive", count = n(), 
              total = nrow(pos_bigrams_separated),
              rel_freq = count/total),
  neg_bigrams_separated %>% 
    filter(word2 == "style") %>%
    summarize(type = "negative", count = n(), 
              total = nrow(neg_bigrams_separated),
              rel_freq = count/total)
)

characters_freq <- bind_rows(
  pos_bigrams_separated %>% 
    filter(word2 == "characters") %>%
    summarize(type = "positive", count = n(), 
              total = nrow(pos_bigrams_separated),
              rel_freq = count/total),
  neg_bigrams_separated %>% 
    filter(word2 == "characters") %>%
    summarize(type = "negative", count = n(), 
              total = nrow(neg_bigrams_separated),
              rel_freq = count/total)
)

print("Relative frequency of 'style':")
print(style_freq)

print("Relative frequency of 'characters':")
print(characters_freq)

```
  Answer:
  B5: Differences Between Positive and Negative Reviews

  For "style," both positive and negative reviews primarily discuss "writing style," but negative reviews mention it almost twice (136 vs. 75 occurrences). Interestingly, the top 5 preceding words remain identical across both review types, suggesting very different, even opposite, opinions can be discussed and conveyed by similar language usage.

  For "characters," negative reviews reference them more specifically, using "these characters" more frequently, while positive reviews uniquely mention "other characters." This suggests negative reviews might focus on criticizing specific character issues, while positive reviews tend to compare characters or discuss their relationships.

  B6: Relative Frequency Insights

  Style appears *slightly* more often in negative reviews (0.00046 vs. 0.00041), but the difference is nominal. However, characters are mentioned substantially more frequently in negative reviews (0.00142 vs. 0.00107) – about 33% higher. This suggests that character flaws may be a stronger driver of negative reviews than stylistic issues, with poorly developed or unlikeable characters being a common criticism when readers dislike a book.


## C: Negation and sentiment

(3 points)

Section 4.1.3 discusses how naive sentiment classification might get the
sentiment wrong of phrases such as 'no fun'. Whereas 'fun' has a
positive score, this bigram as a whole should be considered negative (or
at least should not contribute to the positive score).

Again, do not apply stop word filtering. In this question we will focus
on bigrams with "not" as the first word.

1.  Determine the 5 most frequent words following ‘not’ in positive
    reviews?
    this means the format of [not, xxx]
2.  Annotate word2 with sentiment scores from the AFINN lexicon as in
    the previous assignment (i.e., add a column with sentiment scores
    for the second word).
3.  Create a bar plot as in Figure 4.2 of the textbook.
4.  The plot should show for all the “not . . . ” bigrams: Which 10
    words contribute most in the positive direction? Which 10 words
    contribute the most in the negative direction? (Contribution is
    computed as frequency times sentiment value of the second word).
5.  Repeat steps 1-3 for the negative reviews.
6.  Do you see interesting differences between positive and negative
    reviews? Comment on your findings.

```{r , fig.width=10,fig.height=10}

# C1: Determine the 5 most frequent words following 'not' in positive reviews
pos_not <- pos_bigrams_separated %>%
  filter(word1 == "not") %>%
  count(word2, sort = TRUE) %>%
  head(5)

print("Most frequent words following 'not' in positive reviews:")
print(pos_not)

# C2: Annotate word2 with sentiment scores from AFINN lexicon
AFINN <- get_sentiments("afinn")

pos_not_sentiment <- pos_bigrams_separated %>%
  filter(word1 == "not") %>%
  inner_join(AFINN, by = c(word2 = "word")) %>%
  count(word2, value, sort = TRUE)

# C3 & C4: Bar plot 
pos_contribution <- pos_not_sentiment %>%
  mutate(contribution = n * value) %>% # chapter 2
  arrange(desc(abs(contribution))) %>%
  head(20) %>%  # Get top 20 words by contribution magnitude
  mutate(word2 = reorder(word2, contribution))  # Reorder for visualisation

# Visualisation for positive reviews
ggplot(pos_contribution, aes(contribution, word2, fill = contribution > 0)) +
  geom_col(show.legend = FALSE) +
  labs(x = "Sentiment value × frequency",
       y = NULL,
       title = "Words following 'not' in positive reviews") +
  scale_fill_manual(values = c("darkred", "darkblue"))

# C5: Repeat for negative reviews
neg_not <- neg_bigrams_separated %>%
  filter(word1 == "not") %>%
  count(word2, sort = TRUE) %>%
  head(5)

print("Most frequent words following 'not' in negative reviews:")
print(neg_not)

neg_not_sentiment <- neg_bigrams_separated %>%
  filter(word1 == "not") %>%
  inner_join(AFINN, by = c(word2 = "word")) %>%
  count(word2, value, sort = TRUE)

neg_contribution <- neg_not_sentiment %>%
  mutate(contribution = n * value) %>%
  arrange(desc(abs(contribution))) %>%
  head(20) %>%
  mutate(word2 = reorder(word2, contribution))

# Visualisation for negative reviews
ggplot(neg_contribution, aes(contribution, word2, fill = contribution > 0)) +
  geom_col(show.legend = FALSE) +
  labs(x = "Sentiment value × frequency",
       y = NULL,
       title = "Words following 'not' in negative reviews") +
  scale_fill_manual(values = c("yellow", "darkgreen"))

```

  Answer: C6: Differences Between Positive and Negative Reviews
  In negative reviews, the clear pattern of negating positive qualities can be observed: "not recommend" and "not worth" appear frequently with high contribution.
As a comparison, positive reviews show more nuanced negation patterns. While they do negate negative qualities, for example, "not disappointed", they use negation more moderately overall. The lower contribution magnitudes suggest negation plays a less critical role in positive assessments.

# D: Log Odds

Here we will use the implementation of the log-odds ratio method for
finding keywords as implemented in the tidylo library
(<https://github.com/juliasilge/tidylo>)

The task is to compare keywords for the novels of two authors from the
gutenberg text collection

1.  Select two authors from the gutenberg collection, where the first
    letter of the last name of at least one of the authors corresponds
    to the first letter of your last name. Think of two authors that
    differ in for instance, style, genre, period, or nationality.
    At least one of the authors' name should start with "Z".
    Zane Grey was an American author and dentist. He is known for his popular adventure novels and stories associated with the Western genre.
    Émile Zola was a French novelist, journalist, playwright. More than half of Zola's novels were part of the twenty-volume Les Rougon-Macquart cycle, which details the history of a single family under the reign of Napoléon III. 
    -- quoted from Wikipedia
2.  Download 3-5 texts for both authors (make sure to include the author
    name, will need later). Tokenize the texts and count the words per
    author.
3.  If you have not done so already, combine the two datasets.
4.  Compute the log odds for the word counts per author (see the github
    page of the tidylo library for an example).
5.  Visualise the top-20 highest ranked words for each author in a bar
    plot (following the example if tidylo).
6.  What are the most important differences in keywords between your
    authors? Comment on your findings.

```{r}
# D1: Select authors
# Get metadata
books <- gutenberg_metadata

# Filter for authors with last name starting with 'Z' and English language works
filtered_books <- books %>% 
  filter(
    str_detect(author, "Z"), # Author's last name starts with Z
    language == "en",         # English language works
    has_text == TRUE          # Only books with available text
  )

# Show the first five authors to choose from
head(filtered_books, 30)

```
I choose the authors "Grey, Zane" and "Zola, Émile" for this task. I will download 3 texts for each author.
```{r}
# D2: Download texts
# Download Zane Grey books
# Zane Grey books
grey_book1 <- gutenberg_download(502, meta_fields = "author")
grey_book2 <- gutenberg_download(2057, meta_fields = "author")
grey_book3 <- gutenberg_download(2066, meta_fields = "author")

# Combine Grey books
grey_books <- bind_rows(grey_book1, grey_book2, grey_book3)

# Download Émile Zola books
zola_ids <- c(8511, 8512, 8513)
zola_books <- gutenberg_download(zola_ids, meta_fields = "author")
# For some reason Zane's books do not support this way
# I have to download each book manually and bind them together

# Follow the tidy text format & word count
grey_words <- grey_books %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words) %>%  # Remove stopwords
  count(author, word, sort = TRUE)

zola_words <- zola_books %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words) %>%  # Remove stopwords
  count(author, word, sort = TRUE)

# D3: Combine the data
all_words <- bind_rows(grey_words, zola_words) 

# D4: Calculate log odds
word_log_odds <- all_words %>%
  bind_log_odds(author, word, n) # must be tidy data

# D5: Visualise top 20 words for each author
word_log_odds %>%
  group_by(author) %>%
  slice_max(log_odds_weighted, n = 20) %>% #select the top 20 words with highest log_odds_weighted values
  # already weighted, dont need to do anything, just use
  ungroup() %>%
  # IMPORTANT!! Reorder the words within each author
  #because the same word might appear for both authors
  # but need different positions in each facet
  mutate(word = reorder_within(word, log_odds_weighted, author)) %>%
  # Visualisation
  ggplot(aes(log_odds_weighted, word, fill = author)) +
  geom_col(show.legend = FALSE) +
  # Split the plot into separate panels (facets) for each author
  facet_wrap(~author, scales = "free_y") +
  scale_y_reordered() +
  labs(title = "Most Distinctive Words in Zane vs. Zola",
       x = "Log odds ratio",
       y = NULL)
```

